The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-04-10 manuscript/manuscript.qmd
Add registration timing section

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
date: "2023-04-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international NGOs (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we measure the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations’ issue areas. We argue that the less contensious the issue areas INGOs work in the more likely they are granted greater operational space, measured by thenumber of provinces they are allowed to work in by the registration documentation. China provides an excellent setting for testing the "closing space" and "changing space" theoretical understanding because the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with multilevel Bayesian models. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
bibliography: references.bib
link-citations: true
---

@@

# Load targets 
tar_load(ongo)
tar_load(ongo_mapdata)
tar_load(c(preds_issue, preds_local, preds_timing,
           epreds_issue, epreds_local, epreds_timing))

@@

$$
\begin{aligned}
\text{Count of provinces}\ \sim&\ \operatorname{Ordered\ Beta}(\mu_{i_j}, \phi_y, k_{0_y}, k_{1_y}) & \text{Registered provinces for INGO $i$ in province $j$}\\[8pt]
\mu_{i_j} =&\ (\beta_0 + b_{0_j}) + \beta_1\ \text{Issue area}\ + & \text{Average outcome}\\
&\ \beta_2\ \text{Local connections}\ + \\
&\ \beta_3\ \text{Years since law took effect} \\
b_{0_j} =&\ \mathcal{N}(0, \sigma_0) & \text{Random province offsets} \\[8pt]
\beta_0\ \sim&\ \operatorname{Student\,t}(\nu = 3, \mu = 0, \sigma = 2.5) & \text{Prior for global intercept} \\
\beta_{1..3}\ \sim&\ \mathcal{N}(0, 5) & \text{Prior for global coefficients} \\
\sigma_0\ \sim&\ \operatorname{Student\,t}(3, 0, 2.5) & \text{Prior for between-province variability} \\
\phi_y\ \sim&\ \operatorname{Exponential}(1 / 100) & \text{Prior for within-province variability} \\
k_{0_y}, k_{1_y}\ \sim&\ \operatorname{Dirichlet}(1, 1, 1) & \text{Prior for 0–continuous and continuous–1 cutpoints} \\
\end{aligned}
$$ {#eq-ordbeta}

@@

## Descriptive Results

## Regression Results

@@

Following our first empirical expectation, we predict that INGOs working on less contentious issues should be registered in more provinces and have more geographic flexibility in operating throughout the country. To explore this, we can take advantage of the simultaneous continuous and discrete features of ordered Beta regression and see how often an INGO in each issue area is predicted to work in a single province, across all provinces, or some number in between. @fig-issue-preds-1 presents a summary of predicted posterior outcomes for 1,000 hypothetical INGOs, varying each possible issue area and holding all other variables at their mean or modal values. Several notable trends emerge from the model results. INGOs focused on arts and culture are overwhelming the most likely kind of organization to work nationally—arts and culture organizations are predicted to work in all 32 provinces 80% of the time (with a credible interval of 78–83%) and are only predicted to work in a single province a rare 3–5% of the time. Education-focused INGOs, on the other hand, are the least likely kind of organization to be registered in all 32 provinces and are predicted to do so only 32–38% of the time. Education INGOs are the most likely to be registered in only one province, with that predicted outcome appearing 24–29% of the time—at least 10 percentage points higher than any other issue area's predicted proportion. All other issue areas follow roughly the patterns in predictions: the most common outcome is for organizations to work nationally (50–60% of the time), followed by working in at least 2 provinces (25–40% of the time), followed by working in a single province (10–20% of the time). @fig-issue-preds-2 shows the continuous posterior distribution of the average number of provinces INGOs are expected to be registered in across all types of issue areas, but accounting for the 1- and 32-only provinces simultaneously. We can see a similar trend—arts and culture INGOs are predicted to work in 28–30 provinces on average, while education INGOs are only predicted to work in 12–20. All other issues range between 20–28 provinces.

@@

```{r fig-issue-preds-1, fig.width=6, fig.height=4.5, cache=TRUE}
#| include: true
#| echo: false
#| fig-subcap: "Proportions of posterior predictions of INGOs working in 1 province, all 32 provinces, or somewhere in between. Distinct x-axis categories sum to 100% across the panels."
preds_issue_plot <- preds_issue |> 
  unnest(preds) |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1 province",
    .prediction > 0 & .prediction < 1 ~ "1–32 provinces",
    .prediction == 1 ~ "32 provinces"
  )) |> 
  group_by(outcome, work_field_code1, draw) |> 
  summarize(count = n()) |> 
  group_by(work_field_code1, draw) |> 
  mutate(prop = count / sum(count)) |> 
  ungroup() |> 
  mutate(work_field_code1 = fct_reorder(work_field_code1, prop))

@@

# Discussion

>>>>

---
title: "The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society"
date: "2023-04-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@student.gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international NGOs (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we measure the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations’ issue areas. We argue that the less contensious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are allowed to work in by the registration documentation. China provides an excellent setting for testing the "closing space" and "changing space" theoretical understanding because the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with multilevel Bayesian models. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
bibliography: references.bib
link-citations: true
---

@@

# Load targets 
tar_load(ongo)
tar_load(ongo_mapdata)
tar_load(c(preds_issue, preds_local, preds_timing,
           epreds_issue, epreds_local, epreds_timing,
           epreds_timing_local))

@@

$$
\begin{aligned}
\text{Count of provinces}\ \sim&\ \operatorname{Ordered\,Beta}(\mu_{i_j}, \phi_y, k_{0_y}, k_{1_y}) & \text{Registered provinces for INGO $i$ in province $j$}\\[8pt]
\mu_{i_j} =&\ (\beta_0 + b_{0_j}) + \beta_1\ \text{Issue area}\ + & \text{Average outcome}\\
&\ \beta_2\ \text{Local connections}\ + \\
&\ \beta_3\ \text{Years since law took effect} \\
b_{0_j} =&\ \mathcal{N}(0, \sigma_0) & \text{Random province offsets} \\[8pt]
\beta_0\ \sim&\ \operatorname{Student\,t}(\nu = 3, \mu = 0, \sigma = 2.5) & \text{Prior for global intercept} \\
\beta_{1..3}\ \sim&\ \mathcal{N}(0, 5) & \text{Prior for global coefficients} \\
\sigma_0\ \sim&\ \operatorname{Student\,t}(3, 0, 2.5) & \text{Prior for between-province variability} \\
\phi_y\ \sim&\ \operatorname{Exponential}(1 / 100) & \text{Prior for within-province variability} \\
k_{0_y}, k_{1_y}\ \sim&\ \operatorname{Dirichlet}(1, 1, 1) & \text{Prior for 0–continuous and continuous–1 cutpoints} \\
\end{aligned}
$$ {#eq-ordbeta}

@@

## Descriptive findings

## Posterior predictions

@@

Following our first empirical expectation, we predict that INGOs working on less contentious issues should be registered in more provinces and have more geographic flexibility in operating throughout the country. To explore this, we can take advantage of the simultaneous continuous and discrete features of ordered Beta regression and see how often an INGO in each issue area is predicted to work in a single province, across all provinces, or some number in between. @fig-issue-preds-ordbeta presents a summary of predicted posterior outcomes for 1,000 hypothetical INGOs, varying each possible issue area and holding all other variables at their mean or modal values. Several notable trends emerge from the model results. INGOs focused on arts and culture are overwhelming the most likely kind of organization to work nationally—arts and culture organizations are predicted to work in all 32 provinces 80% of the time (with a credible interval of 78–83%) and are only predicted to work in a single province a rare 3–5% of the time. Education-focused INGOs, on the other hand, are the least likely kind of organization to be registered in all 32 provinces and are predicted to do so only 32–38% of the time. Education INGOs are the most likely to be registered in only one province, with that predicted outcome appearing 24–29% of the time—at least 10 percentage points higher than any other issue area's predicted proportion. All other issue areas follow roughly the patterns in predictions: the most common outcome is for organizations to work nationally (50–60% of the time), followed by working in at least 2 provinces (25–40% of the time), followed by working in a single province (10–20% of the time). @fig-issue-preds-2 shows the continuous posterior distribution of the average number of provinces INGOs are expected to be registered in across all types of issue areas, but accounting for the 1- and 32-only provinces simultaneously. We can see a similar trend—arts and culture INGOs are predicted to work in 28–30 provinces on average, while education INGOs are only predicted to work in 12–20. All other issues range between 20–28 provinces.

@@

```{r fig-issue-preds-ordbeta, fig.width=6, fig.height=4.5, cache=TRUE}
#| include: true
#| echo: false
#| fig-subcap: "Proportions of posterior predictions of INGOs working in 1 province, all 32 provinces, or somewhere in between. Distinct x-axis categories sum to 100% across the panels."
preds_issue_plot <- preds_issue |> 
  unnest(preds) |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1 province",
    .prediction > 0 & .prediction < 1 ~ "1–32 provinces",
    .prediction == 1 ~ "32 provinces"
  )) |> 
  group_by(outcome, work_field_code1, draw) |> 
  summarize(count = n()) |> 
  group_by(work_field_code1, draw) |> 
  mutate(prop = count / sum(count)) |> 
  ungroup() |> 
  mutate(work_field_code1 = fct_reorder(work_field_code1, prop))

@@

The ONGO Law officially came into effect on January 1, 2017, and as seen previously in @fig-registration-timeline, the first year of the law saw the greatest total number of registrations. In our third empirical expectation, we propose that these early registrants under the new law were both less of a threat to regime stability and a higher priority for bureaucratic processing, and thus should be registered in more provinces and have more geographic flexibility throughout the country. However, the results of our model support the *reverse* of this expectation. @fig-timing-preds-1 once again shows the predicted posterior outcomes for 1,000 simulated INGOs across a range of time since January 2017. Organizations that registered in the first year are predicted to work in all provinces 41–46% of the time, and are predicted to work in two or more provinces 34–39% of the time. These early INGOs are the most likely to be registered in only one province and are predicted to do so 17–23% of the time. After the initial wave of registrations in 2017, the likelihood of nationwide reach steadily increases and replaces both the probability of a single province or a partial range of provinces. By the end of 2022—5 years after the ONGO Law took effect—the probability of national -level registration increases to 55–61% while the probability of working in a range of provinces increases to predicted to work in two or more provinces increase to 27–32%. These late INGOs are the least likely to be registered in only one province and are predicted to do so only 17–22% of the time. @fig-timing-preds-2 demonstrates the same results with the posterior distribution of the expected count of provinces over five years since 2017. INGOs that register early are predicted to work in 17–22 provinces, while those that wait until 2022 are predicted to work in 21–28.

::: {#fig-timing-preds layout-ncol=1}

```{r fig-timing-preds-1, fig.width=6, fig.height=3}
#| include: true
#| echo: false
#| fig-subcap: "Proportions of posterior predictions of INGOs working in 1 province, all 32 provinces, or somewhere in between. Distinct x-axis values sum to 100% across the panels."
preds_timing_plot <- preds_timing |> 
  unnest(preds) |> 
  mutate(outcome = case_when(
    .prediction == 0 ~ "1 province",
    .prediction > 0 & .prediction < 1 ~ "1–32 provinces",
    .prediction == 1 ~ "32 provinces"
  )) %>% 
  group_by(outcome, years_since_law, draw) %>% 
  summarize(count = n()) |> 
  group_by(years_since_law, draw) |> 
  mutate(prop = count / sum(count))

# preds_timing_plot |>
#   group_by(outcome, years_since_law) |>
#   median_qi(prop)
# 
# epreds_timing |>
#   mutate(.epred = prop_to_provinces(.epred)) |>
#   group_by(years_since_law) |>
#   median_qi(.epred)

ggplot(preds_timing_plot, aes(x = years_since_law, y = prop)) +
  stat_lineribbon(.width = c(0.5, 0.8, 0.95), 
                  alpha = 0.35, fill = clrs[7], color = clrs[7]) +
  facet_wrap(vars(outcome), nrow = 1) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Years since law took effect (2017)", 
       y = "Proportion of predictions",
       caption = "Shaded regions show the 50%, 80%, and 95% credible intervals") +
  theme_ongo()
```

```{r fig-timing-preds-2, fig.width=6, fig.height=3}
#| include: true
#| echo: false
#| fig-subcap: "Distribution of the expectation of posterior predictions. These distributions take into account single-province and nationwide predictions and represent the overall conditional average geographic range for INGOs."
epreds_timing |> 
  ggplot(aes(x = years_since_law, y = .epred)) +
  stat_lineribbon(.width = c(0.5, 0.8, 0.95), 
                  alpha = 0.35, fill = clrs[7], color = clrs[7]) +
  scale_y_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces",
                                         breaks = seq(15, 27, 3))) +
  guides(fill = "none") +
  labs(x = "Years since law took effect (2017)", y = "Expected proportion of provinces",
       caption = "Shaded regions show the 50%, 80%, and 95% credible intervals") +
  theme_ongo() +
  theme(axis.title.y.right = element_text(hjust = 1, margin = margin(l = 10), angle = 90),
        axis.ticks.y.right = element_line(linewidth = 0.25, color = "grey50"))
```

Posterior predictions from ordered Beta regression model, conditional on time since registration.

We lack sufficient theory or data to explain why this trend goes against our expectations, but we can again make reasonable speculations. As we saw previously with INGOs and local connections, organizations appear to strategically seek out the best avenues and methods for registering. Organizations that participated in the initial wave of registrations in 2017 likely had closer connections with local officials and were able to speed through the bureaucratic process registration more easily. As a preliminary test of this speculation, we rerun our main model with an interaction between years since registration and local connections to see how the effect of local connections changes over time. As seen in @fig-timing-local, early registrants in 2017 without local connections are more likely to have a larger geographic reach, registering in an average of 19–23 provinces, while their locally-focused counterparts are only expected to register in an average of 7–16 provinces. Accordingly, it may be that organizations with local connections are more likely to register early because of their pre-existing bureaucratic relationships, while organizations with a larger national reach take longer to undertake the registration process and thus register later. Further research on this point is warranted.

```{r fig-timing-local, fig.width=6, fig.height=3}
#| include: true
#| echo: false
#| fig-cap: "Distribution of the expectation of posterior predictions across both time and local connections. These distributions take into account single-province and nationwide predictions and represent the overall conditional average geographic range for INGOs."
epreds_timing_local |> 
  mutate(local_connect = factor(local_connect, labels = c("No local connections", "Local connections"))) |> 
  ggplot(aes(x = years_since_law, y = .epred)) +
  stat_lineribbon(.width = c(0.5, 0.8, 0.95), 
                  alpha = 0.35, fill = clrs[7], color = clrs[7]) +
  scale_y_continuous(labels = label_percent(),
                     sec.axis = sec_axis(trans = ~ prop_to_provinces(.),
                                         name = "Expected provinces",
                                         breaks = seq(3, 27, 4))) +
  guides(fill = "none") +
  labs(x = "Years since law took effect (2017)", y = "Expected proportion of provinces",
       caption = "Shaded regions show the 50%, 80%, and 95% credible intervals") +
  facet_wrap(vars(local_connect)) +
  theme_ongo() +
  theme(axis.title.y.right = element_text(hjust = 1, margin = margin(l = 10), angle = 90),
        axis.ticks.y.right = element_line(linewidth = 0.25, color = "grey50"))
```

# Discussion and conclusion
